---
title: "Project 3: stat302package Tutorial"
author: "Ruofeng Tang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stat302package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 1.Introduction

The goal of stat302package is to perform t-test, linear model regression, k-nearest-neighbors algorithm and random forest algorithm on given data. 

You can install the the package from GitHub using:

```{r, eval = FALSE}
devtools::install_github("https://github.com/RuofengT/stat302package")
```

```{r setup}
library(stat302package)
```

## 2.Tutorial for `my_t.test`

This function performs t-test on given data. It takes a numeric vector `x` of data, a a string specifying types of t test wanted with choices `two.sided`, `less` or `greater`, and it takes a number `mu` indicating the null hypothesis. It then conducts t test with given parameters, and returns a list including: the numeric test statistic, the degrees of freedom, the type of t test given, and the numeric p-value.

### Examples

Performing t test on `lifeExp` data from `my_gapminder`, with H_0: mu = 60, H_a: mu ≠ 60:

```{r}

```

## 3.Tutorial for `my_lm`

## 4.Tutorial for `my_knn_cv`

## 5.Tutorial for `my_rf_cv`
